---
title: "Project Proposal"
author: "Data Wranglers"
output: github_document
---

```{r load-packages, echo=F, message = FALSE, warning = F}
library(tidyverse)
```

```{r load-data}
all_stars <- read_csv("../data/all_star.csv", show_col_types = F)
colleges <- read_csv("../data/colleges.csv", show_col_types = F)
states <- read_csv("../data/states.csv", show_col_types = F)
stats <- read_csv("../data/stats.csv", show_col_types = F)
```

### High Level Goal

We want to build a shiny app with visualizations based on NBA All-Star data, including where the athletes are originally from, what colleges they went to, what team they played for, and important statistics from their career. We will answer each of our three research questions described below with separate visualizations in the shiny app, some of which will be interactive.

### Motivation

Our motivation behind our ideas for this project are that we are all NBA fans and thought it was an interesting, unique idea. We found a lot of relevant data to help us create effective and interesting visualizations, which we were able to use by scraping the web. These data are listed below:

### Data

1. NBA All-Stars (1950-2021) - scraped from [Real GM](https://basketball.realgm.com/)

This dataset includes NBA All-Stars from 1950-2021 and information on the team they played for, their draft pick, and where they are from. We will use this dataset for visualizations in the shiny app, including players' nationalities, when they were drafted, and All-Stars by team. The dataset has `r length(all_stars)` variables and `r nrow(all_stars)` observations.


2. NBA Player Birthplaces - scraped from [Basketball Reference](https://www.basketball-reference.com/)

This second dataset has `r length(states)` variables and `r nrow(states)` observations and includes more information on NBA players through a large number of years. The most important column in this dataset is the state from which a player hails, as we will be using this to visualize which state our All-Stars are from; we will join this dataset with the one above to achieve this visualization. The data only includes American born players at present, which is something we may change later.


3. NBA player stats - scraped from [Basketball Reference](https://www.basketball-reference.com/)

This third dataset includes `r length(states)` variables and `r nrow(states)` observation. It contains player statistics in each season played in the NBA since 1951. We will use this dataset to create interactive visualizations of stats and other data.


4. NBA player colleges - scraped from [Basketball Reference](https://www.basketball-reference.com/)

This third dataset includes `r length(colleges)` variables and `r nrow(colleges)` observation. It contains the college that each NBA player has attended. This can be used to show which players produce the most NBA players and NBA All-Stars.


### Research questions

1. Where are NBA All-Stars originally from?

With the NBA becoming an increasingly international league, we are curious to look into where the best players (the ones that made the All-Star game) were originally from. We also thought it would be interesting to create a panel in the Shiny App first demonstrating where NBA players are from globally, and then for those born in the US, what states they were born in. We will join the first dataset, which lists players and their nationalities, to the second dataset, which lists which state players were born in.

2. When were All-Stars drafted and what team did they play for?

For a second question, we will create an animated bar chart where you can manually switch between looking at what teams NBA All-Stars played for and when players were drafted. We chose this because we wanted to see which teams had the most players make the All-Star game in the 2000s, and also were curious how many late draft picks versus earlier draft picks became good enough to make the All-Star game. 

3. Which colleges did NBA players and NBA All-Stars attend?

Our final visualization will show which colleges NBA All-Stars attended on a map and will allow users to search for a college and pull up a list of all the All-Stars from 2000-2016 that attended that college. We chose this question do to our genuine interest in both college basketball and the NBA, as well as wanting to know which colleges produced the most NBA All-Stars.

### Weekly Plan of Attack

Week 1 (week of Mon, Oct 18): 
- Picked theme for project and found relevant datasets that fit our theme - Assigned to entire group


Week 2 (week of Mon, Oct 25): 
- Finalized research questions and worked on our project proposal and weekly plan of attack for the entire project - Assigned to entire team

Week 3 (week of Mon, Nov 1):

- Fix revisions and make correctiosn to finalize proposal - Assigned to entire group

Week 4 (week of Mon, Nov 8): 

- Make final updates to proposal and conduct peer review of other projects - Assigned to entire team
- Combine datasets and start working on different panels of shiny app- Each team member will clean one of the three datasets and will establish one panel of the Shiny App

Week 5 (week of Mon, Nov 15): 

- Each team member will work on one panel for the Shiny app and we will meet weekly (outside of lab) to discuss progress - Sarab will work on the map for "Where All-Stars are originally from?", Zach will work on "When were All-Stars drafted and what team did they play for?", and Owen will work on "Which colleges did NBA All-Stars attend?".

Week 6 (week of Mon, Nov 22): 

- Finish up visualizations and start working on write-up and presentation - Each team member will conduct the appropriate write-ups and interpretations/discussions (along with the slides in the presentation) for their own visualization, we will then appropriately split up the introduction and conclusion for the project as a whole between the 3 of us

Week 7 (week of Mon, Nov 29): 

- Finalize write-up and presentation and final edits
- Conduct peer review - Assigned to entire group


### Repository Organization

1. Data
- Contains files used to scrape data
- Contains .csv files
- Contains data dictionaries for all data files

2. Shiny
- Contains files for work done on Shiny app

3. Presentation
- Contains Shiny App output

4. Proposal
- Contains Proposal
